The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-04-24 paper/ms.tex
found a few more leftover instances of error language in section 2, now fixed

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. We then discuss two potential
sources of these coordinate system offsets in
Sections~\ref{sec:mes_err}~and~\ref{sec:local_file}. In
Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
parameters of the Galactocentric coordinate system. In
Section~\ref{sec:local_fire} we examine the azimuthal variations of the
midplane itself in two sets of simulations: cosmological, hydrodynamical,
zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
simulation of a Sagittarius encounter with a Milky Way-mass galaxy
\citep{2018MNRAS.481..286L}.

@@

\section{Motivation} \label{sec:ref_frame}
We first demonstrate the significance of systematic errors in the
determination of the Galactic midplane and distance to the Galactic center to
action computations. We will find that such errors are especially important
for disk-like orbits. The consequences we explore here may also arise from
various other systematic errors. For instance, the axisymmetric Galactic
potential model used in many works to compute actions may not be a good
description of the true potential --- or the parameters used may yield a
potential that is systematically incorrect outside an original fitted region.
In this work, we assume that the Galaxy is perfectly described by our model
axisymmetric potential, and simply explore the consequences of systematic
errors in the determination of the Galactocentric coordinate system.

\subsection{Cartoon Effect of Midplane Error} \label{ssec:cartoon}
We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
inaccurate determination of the midplane leads to a phase-dependence of the
actions calculated from an orbit specified by a phase-space point and an
assumed potential model. The $y$-axis corresponds to the vertical height of
the orbit and the $x$-axis orbital phase. The solid gray curve indicates the
``true'' orbit of the star; that is, its orbit as it oscillates around the
true midplane. The dashed gray line is an erroneous midplane assigned to the
potential model which an observer uses to integrate the orbit of the star and
hence estimate its actions. The model potential is otherwise identical to the
one in which the star is actually moving.

Now suppose this observer makes a measurement of the true orbit at the purple
point or the green point (i.e. at two different orbital phases). Then the
purple and green curves correspond to the orbits that the observer would
compute for each point based on the potential model with the erroneous
midplane. In action space, this would correspond to a different value of $J_z$
for the purple and green points. In this way, assuming the wrong coordinate
system induces a phase-dependence in the actions estimated for the star, which
in the correct potential (in this example, the one with the correct midplane)
should be phase-independent.

>>>>

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. We then discuss two potential
sources of these coordinate system offsets in
Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
parameters of the Galactocentric coordinate system. In
Section~\ref{sec:local_fire} we examine the azimuthal variations of the
midplane itself in two sets of simulations: cosmological, hydrodynamical,
zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
simulation of a Sagittarius encounter with a Milky Way-mass galaxy
\citep{2018MNRAS.481..286L}.

@@

\section{Motivation} \label{sec:ref_frame}
We first demonstrate the significance  to
action computations of a systematic offset in the
determination of the Galactic midplane or distance to the Galactic center. We will find that such offsets are especially important
for disk-like orbits. The consequences we explore here may also arise from
various other systematic errors. For instance, the axisymmetric Galactic
potential model used in many works to compute actions may not be a good
description of the true potential --- or the parameters used may yield a
potential that is systematically incorrect outside an original fitted region.
In this work, we assume that the Galaxy is perfectly described by our model
axisymmetric potential, and simply explore the consequences of offsets in the determination of the Galactocentric coordinate system.

\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
inaccurate or erroneous determination of the midplane (and hence the zero-point of the $z$ coordinate) leads to a phase-dependence of the
actions calculated from an orbit specified by a phase-space point and an
assumed potential model. The $y$-axis corresponds to the vertical height of
the orbit and the $x$-axis shows orbital phase. The solid gray curve indicates the
``true'' orbit of the star; that is, its orbit as it oscillates around the
true midplane. The dashed gray line is the midplane location in the
potential model used to integrate the orbit of the star and
hence estimate its actions, which is offset from the true midplane of the star's orbit. The model potential is otherwise identical to the one in which the star is actually moving.

Now suppose this observer makes a measurement of the star's position and velocity at the purple
point or the green point (i.e. at two different orbital phases). Then the
purple and green curves correspond to the orbits that the observer would
compute for each point based on the potential model with the offset
midplane. In action space, this would correspond to a different value of $J_z$
for the purple and green points. In this way, assuming the wrong coordinate
system induces a phase-dependence in the actions estimated for the star, which
in the correct potential (in this example, the one with the correct midplane)
should be phase-independent.
